---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

[![CRAN Status](https://www.r-pkg.org/badges/version/MetaStan)](https://cran.r-project.org/package=MetaStan) 
[![Total Downloads](http://cranlogs.r-pkg.org/badges/grand-total/MetaStan)](http://cranlogs.r-pkg.org/badges/grand-total/MetaStan)


MetaStan R package
==================

`MetaStan` is an R package for Meta-Analysis using Stan.


Building
--------
`MetaStan` package is on CRAN. `MetaStan` can be installed as follows:

```{r, echo=TRUE, eval=FALSE}
install.packages("MetaStan")
```


The development version of the package can be installed using `devtools`:

```{r, echo=TRUE, eval=FALSE}
devtools::install_github("gunhanb/MetaStan")
```


Vignette
--------
The vignette can be obtained by:
```{r, echo=TRUE, eval=FALSE}
vignette("MetaStan_BNHM")
```


Testing
-------

The `testthat` package is used for testing. Tests reside in the
`tests/testthat` directory. 


Licence
-------

The MetaStan package was written by [Burak Kürsad Günhan](http://bkguenhan.rbind.io). It is licensed under the [GNU General Public License](http://www.gnu.org/licenses/old-licenses/gpl-2.0.txt). For citation info, type `citation(package='MetaStan')` in R. To report any issues or bugs or to suggest enhancements to the package, please go [here](https://github.com/gunhanb/MetaStan/issues).
